---
title: "Class 4- Data visualization"
description: |
  Ggplot2: The language of graphs.
author:
  - name: Eveline Miranda
    url: {}
date: 2021-12-16
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```

## Aesthetic Mappings and Facets


```{r cars}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()

```

## Including Plots

You can also embed plots, for example:

```{r pressure1, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Regression scatterplot    
*  Highway miles per gallon on displacement  
*  Color by drive train (drv)  
*  Linetype by drive train  
*  Linear regression with no confidence interval  
*  Increase the size of the lines  

```{r linear regression1}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(mapping = aes(linetype = drv), method = "lm", se = FALSE, size = 2)
```

## Loess smoothed scatterplot with points of various sizes
*  Highway miles per gallon on displacement    
*  Color points by car class  
*  Point size by city miles per gallon (cty)  
*  Large red loess line for all point and no confidence interval  

```{r linear regression for all pointsA, message = FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class, size = cty)) +
  geom_smooth(se = FALSE, size = 2, color = "red")
```

## Linear regression of highway on displacement
*  Color by vehicle class
*  Simple regression by vehicle class
*  Show all points

```{r regression by vehicle classa}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm")
```

## Linear regression of highway on displacement
*  Color by vehicle class
*  Simple regression by vehicle class
*  Each class has its own scatterplot
*  Show all the points  
*  No Legend

```{r regression by vehicle class with separate scatterplotsa}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point(position = "jitter", show.legend = FALSE) +
  geom_smooth(method = "lm", show.legend = FALSE) +
  facet_wrap(vars(class))
```  

## Linear regression of highway on displacement
*  Color by drivetrain type
*  Simple regression (hwy on displacement) by drive train type and number of cylinders
*  Draw a larger, black regression line
*  Each drive train and number of cylinders has its own scatterplot
*  Show all the points  
*  No Legend

```{r regression by vehicle class with drive train by cylinders A}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point(position = "jitter", show.legend = FALSE) +
  geom_smooth(method = "lm", show.legend = FALSE, color = "black", size = 1) +
  facet_grid(rows = vars(as.factor(cyl)), cols = vars(drv))
```  

## Bar chart of car classes  
*  Bar colored by car class  
*  No legend  

```{r bar chart by car class-a}
ggplot(data = mpg, mapping = aes(x = class, fill = class)) +
  geom_bar(show.legend = FALSE)
```

## Bar chart of car classes  
*  Bar colored by car class  
*  Stack bar in each by drive type  
*  Proportion of drive for each class of vehicle  

```{r bar chart by car class and drive type}
ggplot(data = mpg, mapping = aes(x = class, fill = drv)) +
  geom_bar(position = "fill")
```  

## Bar chart of car classes  
*  Vertical orientation  
*  Number of by drive train type for each vehicle class  

```{r bar chart by car class and drive type number of vehicles}
ggplot(data = mpg, mapping = aes(x = class, fill = drv)) +
  geom_bar(position = "dodge")
```  

```{r geom bar}
ggplot(data = mpg, mapping = aes(x = class, fill = class)) +
  geom_bar()
```

##Columns
 group_by
 Summarize 
 class_count

```{r columns}
class_count <- mpg %>% 
  group_by(class) %>% 
  summarize(count = n())
class_count %>% print()
```

##Distribution numerical variables

How to summarize numerical data?

Histograms
```{r  histograms}
ggplot(data = mpg, mapping = aes(x = hwy)) +
  geom_histogram(fill = "blue", binwidth = 2.5)
```

Frequency polygons
```{r frequency polygons}
ggplot(data = mpg, mapping = aes(x = hwy)) +
  geom_freqpoly(color = "blue", binwidth = 2.5)
```

Densities

``` {r Densities}
ggplot(data = mpg, mapping = aes(x = hwy)) +
  geom_density(fill = "blue")
```

RUG

```{r rug}
ggplot(data = faithful, mapping = aes(x = eruptions)) +
  geom_rug()
```

##Combinations of density plots

``` {r combination  of densities plots}
ggplot(data = faithful, mapping = aes(x = eruptions)) +
  geom_histogram(fill = "lightblue")
```

```{r 1}
ggplot(data = faithful, mapping = aes(x = eruptions)) +
  geom_freqpoly(color = "lightblue")
```

```{r 3}
ggplot(data = faithful, mapping = aes(x = eruptions)) +
  geom_freqpoly(color = "lightblue")
```

```{r 4}
ggplot(data = faithful, mapping = aes(x = eruptions)) +
  geom_density(fill = "lightblue")
```

##combinations:
```{r combined}
ggplot(data = faithful, mapping = aes(x = eruptions)) +
  geom_histogram(aes(y = stat(density)), fill = "lightgreen") +
  geom_density(fill = "blue", alpha = 0.3) +
  geom_rug(color = "red")
```

# Violin Plots

```{r violin 1}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, fill = class)) +
  geom_violin(show.legend = FALSE)
```

# Boxplot
```{r boxplot}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, fill = class)) +
  geom_boxplot(show.legend = FALSE)
```

# Violin plots:
```{r violin plots}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, fill = class)) +
  geom_violin(show.legend = FALSE) +
  geom_boxplot(show.legend = FALSE, width = 0.05, fill = "grey")
```

# Numerical variables

##Scatterplot

```{r scatterplot}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() 
```

## Aesthetic Mappings and Facets

