---
title: "Class 4- Data visualization"
description: |
  Ggplot2: The language of graphs.
author:
  - name: Eveline Miranda
    url: {}
date: 2021-12-16
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(tidyverse)
```

Using an R dataset
```{r data}
data(mpg)
?mpg
glimpse(mpg)
summary(mpg)
```
# Looking at mpg$year
```{r  class}
summary(mpg$year)

# Change year to factor
mpg$year<- as.factor(mpg$year)

# class of mpg$year
class(mpg$year)

# another summary
summary(mpg$year)
```
#look at mpg$trans
```{r Changing class}
summary(mpg$trans)
mpg$trans<-as.factor(mpg$trans)
summary(mpg$trans)
summary(mpg$cyl)
plot(mpg$cyl)
plot(mpg$trans)
mpg$displ<-as.double(mpg$displ)
```
#Time for playing with variables and changing those that are needed

# ggplot
```{r data visualization 1}

# ggplot 
ggplot(mpg, mapping = aes(x= displ, y = hwy)) +
  geom_point()
```

```{r data visualization 2}
ggplot(mpg, mapping = aes(x= displ, y = hwy)) +
  geom_point()+
  geom_smooth()
```

```{r data visualization 3}
ggplot(mpg, mapping = aes(x= displ, y = hwy)) +
  geom_point()+
  geom_smooth(method="lm", se=FALSE)
```

```{r data visualization 4}
ggplot(mpg, mapping = aes(x= displ, y = hwy)) +
  geom_point()+
  geom_smooth(method = "lm")
```

```{r data visualization 5}
ggplot(mpg, mapping = aes(x= displ, y = hwy)) +
  geom_point(col="blue")+
  geom_smooth( se = FALSE, col="red")+
  theme_bw()+
  facet_grid(vars(year))
```

```{r data visualization 6}
ggplot(mpg, mapping = aes(x= displ, y = hwy)) +
  geom_point()+
  geom_smooth(method = lm, se = FALSE)+
  geom_smooth(color = "red")
```

```{r data visualization 7}
ggplot(mpg, mapping = aes(x= displ, y = hwy)) +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  geom_smooth(color = "red", se = F)+
  theme_bw()
```

```{r data visualization 8}
ggplot(mpg, mapping = aes(x= displ, y = hwy, color = class)) +
  geom_point()+
  geom_smooth(method = "lm", se = F)
  
```
#class 2- DATA VISUALIZATION- geoms + mapping and facets - good day for practicing- students will delete, change configurations and see the results
##teach summary and how to use help in R
## teach how to use ggplot assist install.packages("ggThemeAssist") - select the code and go to addins - I added addins using view show TOOLBAR
## how to use HELP and the cheat sheet

```{r data visualization 9}
summary(mpg$hwy)
ggplot(data= mpg, mapping = aes(x= displ, y= hwy))+
  geom_point(aes(size=drv))+
  geom_smooth(mapping = aes(linetype= drv), method ="lm", se = F, size= 1)
```

```{r data visualization 10}
ggplot(data= mpg, mapping = aes(x= displ, y= hwy))+
  geom_point(aes(size=class, color= drv))+
  geom_smooth(mapping = aes(linetype= drv), method ="lm", se = F, size= 1)
```

```{r data visualization 11}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
  geom_point(position = "jitter", show.legend = F) +
  geom_smooth(method = "lm", se= F, show.legend = T)+
  facet_wrap(vars(class), ncol=2)
```

```{r data visualization 12}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point(position = "jitter", show.legend = FALSE) +
  geom_smooth(method = "lm", show.legend = FALSE, color = "black", size = 0.5, se= F) +
  facet_grid(rows = vars(as.factor(cyl)), cols = vars(drv))
```

```{r data visualization 13}
ggplot(data = mpg,
       mapping = aes(x = displ, y = hwy, group = drv, color = drv)) +
  geom_point(size = 2, show.legend = T) +
  geom_smooth(color = "blue", se = FALSE)
```

```{r data visualization 14}

ggplot(data = mpg,
       mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point(mapping = aes(color = drv), size = 2) +
  geom_smooth(se = FALSE, color = "blue")
```
#not writing data= and mapping

```{r data visualization 15}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point( aes(color = drv), size = 2) +
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)
```

```{r data visualization 16}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, colour = year)) +
  geom_point(
    size = .2,
    shape = 7,
    stroke = 1, 
    alpha=.4
  )+
  theme_minimal()

```

