---
title: "Class 3- Data Manipulation II"
description: |
  Tidyverse and more
author:
  - name: Eveline Miranda
    url: {}
date: 2021-12-16
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```

##Columns
 group_by
 Summarize 
 class_count

```{r columns}
class_count <- mpg %>% 
  group_by(class) %>% 
  summarize(count = n())
class_count %>% print()
```

# Filtering and Arranging

```{r glimpse}
data(flights)
glimpse(flights)
```



```{r summary}

summary(flights)
```

## Head and Tail
```{r head and tail}
head(flights)
tail(flights)
```


```{r 30 minuts}
make_up <- filter(flights,
                  dep_delay >= 60,
                  dep_delay - arr_delay > 30)
```

Top 10 most departure delayed flights

```{r top 10}
delayed <- arrange(flights, desc(dep_delay))
head(delayed, 20)

```

shortest flight

```{r shortest}
head(arrange(flights, distance), 1)

head(arrange(flights, air_time), 1)
```

```{r arrange}
flights %>%
  arrange(distance) %>%
  head(1)
```


```{r arr}
flights %>% 
  arrange(air_time) %>% 
  head(1)
```

Longest flight
```{r longest}

flights %>%
  arrange(desc(distance)) %>%
  head(1)
```

```{r distance}
flights %>% 
  arrange(desc(air_time)) %>% 
  head(1)
```


#Select and mutation

Give multiple ways to select dep_time, dep_delay, arr_time, and arr_delay from the flights data frame.

```{r select}
variables <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
departure_delays <- select(flights, variables)
departures_delays <-
  select(flights, dep_time, dep_delay, arr_time, arr_delay)
departures_delays <-
  select(flights, starts_with("dep"), starts_with("arr"))
departures_delays <-
  select(flights, ends_with("time"), ends_with("delay"))
departures_delays <- select(flights, 4, 6:7, 9)
```

Calculate the difference between dep_time and sched_dep_time and compare the result with dep_delay. How do we explain the difference?

```{r transmute}
departure_comparison <-
  transmute(flights,
            dep_time,
            sched_dep_time,
            dep_delay,
            time_difference = dep_time - sched_dep_time,
            same = (dep_delay == time_difference))
head(departure_comparison, 20)
```


#Rank

```{r rank}
largest <- mutate(diamonds,
                  size_rank = min_rank(desc(carat)),
                  price_rank = min_rank(desc(price)))

largest <- diamonds %>%
  mutate(size_rank = min_rank(desc(carat)),
         price_rank = min_rank(desc(price)))

# Largest by size
largest %>%
  arrange(size_rank) %>%
  select(carat, size_rank)
```



```{r largest}
largest %>%
  arrange(price_rank) %>%
  select(price, price_rank)
```

# Grouped Summaries


```{r avarage delay}

summarise(
  flights,
  mean_delay = mean(dep_delay, na.rm = TRUE),
  median_delay = median(dep_delay, na.rm = TRUE)
)
```

Average and median delay

```{r Average and median delay}
by_airport <- group_by(flights, origin)
summarize(
  by_airport,
  mean_delay = mean(dep_delay, na.rm = TRUE),
  median_delay = median(dep_delay, na.rm = TRUE)
)
```

average distance
```{r average distance}
by_dest <- group_by(flights, dest)
delay <- summarise(
  by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay
```

average and median delay
```{r average and median delay}
flights %>%
  group_by(origin) %>%
  summarize(
    mean_delay = mean(dep_delay, na.rm = TRUE),
    median_delay = median(dep_delay, na.rm = TRUE)
  )
```

#Filter, group by and summarize

```{r filter and group by}
flights %>%
  filter(dest %in% c("ATL", "ORD", "LAX")) %>%
  filter(carrier %in% c("AA", "DL", "UA")) %>%
  group_by(origin, carrier, dest) %>%
  summarize(
    number_flights = n(),
    location = mean(arr_delay, na.rm = TRUE),
    scale = sd(arr_delay, na.rm = TRUE)
  )
```





Not cancelled flights
```{r not cancelled flights}
not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))


not_cancelled %>% 
  group_by(tailnum) %>% 
  summarize(
    delay = mean(arr_delay),
    n = n()
  )

```



```{r total}
not_cancelled %>% 
  group_by(origin) %>% 
  count(carrier)
```

```{r before }

not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r not.}


not_cancelled %>% 
  group_by(tailnum) %>% 
  summarize(
    delay = mean(arr_delay),
    n = n()
  )

not_cancelled %>% 
  filter(dep_time < 500) %>% 
  count(month, day)


not_cancelled %>% 
  group_by(month, day) %>% 
  summarize(n_early = sum(dep_time < 500))


not_cancelled %>% 
  group_by(origin) %>% 
  summarize(proportion = mean(dep_delay > 60))
```

```{r select and filter }

flights %>% 
  group_by(origin) %>%
  mutate(departure_rank = min_rank(desc(dep_delay))) %>% 
  filter(departure_rank == 1) %>% 
  select(origin, carrier, dep_delay, month, day, sched_dep_time)

```


```{r table X}
data(diamonds)
summary(diamonds)

```

#Exploratory Data Analysis
Formulate questions about your data
Investigate answers to questions by visualizing, transforming, and modelling
Evolve and spawn new questions based on what is learned
Data cleaning also involves visualization, transformation, and modelling.

#Questions
"The key to asking quality questions is to generate a large quantity of questions."

##Two types of questions:

How is a variable distributed?
How are variables related to each other?
Basic definitions
Variable: quantity, quality, or property
Value: state or level of a variable when it is measured
Observation: "a set of measurements under similar conditions"
Tabular data: "a set of values, each associated with a variable and an observation."
##Distribution
Possible values
Frequency or probability of observing values
Describing the distribution gives us a basic set of questions to start asking. These in turn provoke additional questions. When considering a distribution, we should organize our preliminary investigation around:

##Location or central tendency (mean and median)
Scale or spread (sd, range, and interquartile range)
Symmetry
Outliers

#Quantiles
```{r quantiles}

quantile(diamonds$carat)
summary(diamonds$carat)

```

#Interquartile range
The length of the box in the boxplot is called the interquartile range. It is the third quartile minus the first quartile.


#Covariation

```{r}
summary(mpg$hwy)
sd(mpg$hwy)
IQR(mpg$hwy)
```

```{r table 1}
set.seed(1234) # Needed so all get the sample random choice of observations
diamonds2 <- diamonds %>%
  sample_n(1000)

```

```{r table 2}
diamonds2 %>%
  count(clarity)

```
#Linear Model
```{r table 3}
model <- lm(log(price) ~ log(carat) + clarity + color + cut + cut * log(carat),
            data = diamonds2)
model %>% summary()

```

```{r table 4}
model %>% anova()

```

```{r table 5}
# Load workspace
Movies <- read_csv(url("http://govfinance.byu.edu/govfinance/classes/dataScience/lectures/illustrations/Movies.csv"))

Movies %>% glimpse()
```
